---
title: "Advanced Example"
output:
  github_document:
    preserve_yaml: FALSE
vignette: >
  %\VignetteIndexEntry{Advanced Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(fig.path = "figures/", echo = TRUE, warning = FALSE, message = FALSE, comment = NA)
devtools::source_gist("8e6e5dc401e3fc1042ef7a030f9d19c7", filename = "revised_toc.R")
```

```{r setup, include=FALSE}
source("http://raw.githubusercontent.com/cwendorf/dsitR/main/source-dsitR.R")
```

## Advanced Example

This vignette provides a more advanced example of how to use the `dsitR` package to conduct a simulation, visualize the results, and assess the outcomes. The example is designed to be more complex and detailed, making it suitable for those who have some experience with the package or with agent-based modeling in general. It uses force-directed graphs to visualize the simulation, which can reveal more complex patterns and interactions among agents. The assessment section includes a Principal Components Analysis (PCA) to identify underlying dimensions of opinion change and clustering to identify groups of agents with similar opinion trajectories.

```{r toc, echo=FALSE}
knitr::current_input() |> revised_toc(base_level = 3, toc_depth = 1)
```

### Conduct the Simulation

Create the agents and neighborhoods for the network.

```{r}
# Initialize a larger network with multiple opinions
agents <- create_agents(rows=10, cols=10, opinions=4)

# Use a more complex neighborhood structure
neighborhood <- create_neighborhood(agents, neighbors_moore_outside)
```

Run and verify the simulation based on the neighborhood.

```{r}
# Run a longer and slower simulation
result <- run_simulation(agents, neighborhood=neighborhood, steps=100, rate=.05)

# View the first few rows of the simulation to see what is produced
head(result)
```

### Visualize the Simulation

Visualize the a step using a force-directed graph.

```{r, advanced-network}
# Visualize just the initial step
frame1 <- subset(result, time==1)
plot_network(frame1,
  neighborhood=neighborhood,
  opinion="opinion1",
  main="Initial Force-Directed Network"
)
```

Visualize the entire simulation to unveil changes and patterns.

```{r, advanced-animated, eval=FALSE}
# Animate the changes in the force-directed graph
animate_network(result,
  neighborhood=neighborhood,
  opinion="opinion1",
)
```

### Assess the Simulation

Prepare for and conduct the Principal Components extraction.

```{r}
# Extract opinions from the final frame
frame <- subset(result, time == 100)
frame <- frame[, c("opinion1", "opinion2", "opinion3", "opinion4")]
head(frame)

# Run PCA and calculate component scores
frame_scaled <- scale(frame)
pca <- prcomp(frame_scaled, center = TRUE, scale. = TRUE)
scores <- pca$x
```

Determine the number of likely clusters in the data.

```{r, advanced-pca}
# k-means clustering in component space
summary(pca)
pca$rotation
var_explained <- summary(pca)$importance[2,]
plot(var_explained, type="b", pch=19,
     xlab="Principal Component",
     ylab="Proportion of Variance Explained",
     main="Scree Plot")

# Choose number of clusters for later analyses
k <- 2
clusters <- kmeans(scores[,1:2], centers = k)$cluster
```

Assess the difference in component scores by cluster.

```{r, advanced-clusters}
# Boxplots of components per cluster
boxplot(scores[,1] ~ clusters,
        xlab = "Cluster", ylab = "PC1",
        main = "PC1 distribution per cluster",
        col = rainbow(k))
boxplot(scores[,2] ~ clusters,
        xlab = "Cluster", ylab = "PC2",
        main = "PC2 distribution per cluster",
        col = rainbow(k))

# Scatterplot in PC1-PC2 space, colored by clusters
plot(scores[,1], scores[,2],
     col = clusters,
     pch = 19, cex = 1.2,
     xlab = "PC1", ylab = "PC2",
     main = "Agents in PCA space with clusters")
legend("topright", legend = paste("Cluster", 1:k), col=1:k, pch=19)
```
