---
title: "Simple Example"
output:
  github_document:
    preserve_yaml: FALSE
vignette: >
  %\VignetteIndexEntry{Simple Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(fig.path = "figures/", echo = TRUE, warning = FALSE, message = FALSE, comment = NA)
devtools::source_gist("8e6e5dc401e3fc1042ef7a030f9d19c7", filename = "revised_toc.R")
```

```{r setup, include=FALSE}
source("http://raw.githubusercontent.com/cwendorf/dsitR/main/source-dsitR.R")
```

## Simple Example

This vignette provides a simple example of how to use the `dsitR` package to conduct a simulation, visualize the results, and assess the outcomes. The example is designed to be straightforward and easy to follow, making it suitable for those new to the package or to agent-based modeling in general. It uses grid plots to visualize the simulation, making it easy to see the changes in opinions over time.

```{r toc, echo=FALSE}
knitr::current_input() |> revised_toc(base_level = 3, toc_depth = 1)
```

---

### Conduct the Simulation

Identity the simulation parameters to determine the size and length of the simulation.

```{r}
rows <- 8
cols <- 8
opinions <- 2
steps <- 50
rate <- 0.2
```

Create the neighborhood based on the parameters and the type of influence network.

```{r}
# Initialize agents and initial opinions
agents <- create_agents(rows=rows, cols=cols, opinions=opinions)

# Define the interactions among the agents
neighborhood <- create_neighborhood(agents, neighbors_moore)
```

Run and verify the simulation based on the neighborhood and parameters.

```{r}
# Run simulation using the neighborhood
result <- run_simulation(agents, neighborhood=neighborhood, steps=steps, rate=rate)

# View the first few rows of the simulation to see what is produced
head(result)
```

### Visualize the Simulation

Plot the neighborhood at specific steps (either with or without the connections among agents).

```{r, simple-network}
# Plot Opinion 1 at the first step with edges
plot_step(
  subset(result, time == 1),
  neighborhood = neighborhood,
  opinion = "opinion1",
  edges = TRUE
)

# Plot Opinion 1 at the final step without edges
plot_step(
  subset(result, time == 50),
  neighborhood = neighborhood,
  opinion = "opinion1",
  edges = FALSE
)
```

Plot individuals at specific steps to highlight who they are influencing (and are influenced by).

```{r, simple-individual}
# Look at agent 42’s neighborhood at first step
frame <- subset(result, time == 1)
plot_individual(frame, neighborhood, id=42, opinion="opinion1")

# Look at agent 42’s neighborhood at final step
frame <- subset(result, time == 50)
plot_individual(frame, neighborhood, coords=c(2,6), opinion="opinion1")
```

Visualize the entire Simulation to unveil changes and patterns (with or without connections).

```{r, simple-animated, eval=FALSE}
# Animate all steps (with edges) for Opinion 1
animate_simulation(
  result,
  neighborhood = neighborhood,
  opinion = "opinion1",
  edges = TRUE
)

# Animate all steps (without edges) for Opinion 2
animate_simulation(
  result,
  neighborhood = neighborhood,
  opinion = "opinion2",
  edges = FALSE
)
```

### Assess the Simulation

Calculate metrics for specific steps in the simulation.

```{r}
# Calculate the metrics for the first step
frame <- subset(result, time == 1)
compute_metrics(frame)

# Calculate the metrics at the final step
frame <- subset(result, time == 50)
compute_metrics(frame)
```

Calculate and plot trends in the metrics.

```{r simple-metrics}
# Calculate the trends in the metrics
trends <- compute_trends(result)
names(trends)

# Plot the chosen metrics separately for opinions
plot_trend(trends, metric="correlation", opinion="opinion1_opinion2")
plot_trend(trends, metric="consolidation", opinion="opinion1")
plot_trend(trends, metric="diversity", opinion="opinion2")
```
